---
title: "Calling Clojure from R"
output: md_document
---

In this tutorial we show how to call Clojure from R using Renjin.

Renjin, when used [as an R package](http://docs.renjin.org/en/latest/package/), allows to run R code on that JVM and pass data to/from it, for most of the main R datatypes.

## Demo

### Setup

```{r}
library(magrittr) # for the %>% "pipe" operator
options(java.parameters = c("-Xmx1g")) # Set your JVM options before running the JVM
library(rJava)
library(renjin) # Loading this library actually runs the JVM already.

# Add the Clojuress JAR to the classpath.
.jaddClassPath("../target/clojuress-0.1.0-SNAPSHOT-standalone.jar")
```

Is there anything missing in the classpath?

```{r}
jars_in_classpath <- renjin({
  import(java.lang.Thread)
  import(java.util.Arrays)
  Arrays$asList(Thread$currentThread()$getContextClassLoader()$getURLs()) %>%
    sapply(as.character)
}) %>%
  gsub(pattern="file:", replacement="")

jars_in_classpath[!file.exists(jars_in_classpath)]
```


### Running code

The following function will allow us to run Clojure code.

We define it using a Renjin block, that runs on the JVM and covnerts the returned value from a Renjin object to an R object. For more details about this way of using Renjin, look [here](http://docs.renjin.org/en/latest/package/index.html) at the Renjin docs.


```{r}
## This function will be run on the JVM through Renjin.
clj_impl <- function(code) {
  import(clojure.java.api.Clojure)
  ## Given a qualified name of a clojure var, we can get its value.
  clj_get <- (function(name) Clojure$var(name))
  ## Given an unqualified name, let us take it from the clojure.core namespace:
  clj_core_get <- (function(name) Clojure$var("clojure.core", name))
  ## Let us get some Clojure functions (as Java objects implementing the interface IFn):
  clj_eval <- clj_core_get("eval")
  require <- clj_core_get("require")
  read_string <- clj_core_get("read-string")
  comp <- clj_core_get("comp")
  ## IFn has a method named 'invoke', that can be used in renjin as a function.
  ##
  ## For example, let us invoke the clojure function 'comp':
  eval_read_string <-
    comp$invoke(
      clj_eval,
      read_string)
  ## Using 'comp', we just composed two Clojure functions and got a new one, that:
  ## It reads a given string of clojure code,
  ## evaluates the corresponding form
  ## and returns the result.
  ##
  ## Let us use this function to require a namespace from Clojuress.
  eval_read_string$invoke("(require 'clojuress.renjin.to-renjin)")
  ## Now we may use functions of these namespace:
  to_renjin <- clj_get("clojuress.renjin.to-renjin/->renjin")
  ## Let us create yet another function composition:
  eval_read_string_to_renjin <-
    comp$invoke(
      to_renjin,
      clj_eval,
      read_string)
  ## We just created a Clojure function that does the following:
  ## Reads a given string of clojure code,
  ## evaluates the corresponding form,
  ## converts the result to a renjin object
  ## and returns that object.
  ##
  ## Let us use this function to read, evaluate and convert the code we got:
  eval_read_string_to_renjin$invoke(code)
  ## The returned value of this block is the result of this evaluated.
  # It will be converted to an R object, if Renjin knows how.
  # For many datatypes, Renjin does know how.
}

#' @export
clj <- function(code) renjin(clj_impl(code))
```

Let us try it:

```{r}
clj("(+ 1 2)")
```

```{r}
clj("{:x [1 nil 2] :y [:A :B]}")
```

### Using functions

The following function will allow us to use a clojure function as an R function. The R arguments are be converted to a Clojure data structure, and the return value is covnerted back from Clojure to R.

```{r}
clj_fn_impl <-
  function(fn_code) function(args) {
    import(clojure.java.api.Clojure)
    clj_get <- (function(name) Clojure$var(name))
    clj_core_get <- (function(name) Clojure$var("clojure.core", name))
    clj_eval <- clj_core_get("eval")
    require <- clj_core_get("require")
    read_string <- clj_core_get("read-string")
    comp <- clj_core_get("comp")
    eval_read_string <-
      comp$invoke(
        clj_eval,
        read_string)
    eval_read_string$invoke("(require 'clojuress.renjin.to-renjin)")
    to_renjin <- clj_get("clojuress.renjin.to-renjin/->renjin")
    eval_read_string$invoke("(require 'clojuress.renjin.to-clj)")
    to_clj <- clj_get("clojuress.renjin.to-clj/->clj")
    eval_read_string_to_renjin <-
      comp$invoke(
        clj_eval,
        read_string)
    fn_from_code <- eval_read_string$invoke(fn_code)
    f <- comp$invoke(
      to_renjin,
      fn_from_code,
      to_clj)
    f$invoke(args)
  }

#' @export
clj_fn <- function(fn_code) function(args) renjin(clj_fn_impl(fn_code)(args))
```

Let us use it:

```{r}
clj_fn("(partial map (partial * 10))")(1:9)
```

```{r}
clj_fn("(partial group-by even?)")(1:9)
```

```{r}
clj("(require '[com.rpl.specter :refer [transform ALL MAP-VALS]]) nil")
clj_fn("#(identity (transform [ALL MAP-VALS ALL] (partial * 100) %))")(
    list(list(x=1:3,
              y=(1:3)^2),
         list(x=1:4,
              y=(1:4)^2)))
```

## Discussion

TODO: Explain in detail the data conversion flow in the above examples.
